Grove: A Convergent Collaborative Structure-Editor Calculus

### 2020-09-30 grove-paper.tex
Commit further work on figures

%% Packages
\usepackage{booktabs}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{semantic}
\let\colonapprox\undefined % Avoid redefinition error in `colonequals`
\let\colonsim\undefined % Avoid redefinition error in `colonequals`
\usepackage{colonequals}
\usepackage{mathpartir}
\usepackage{stmaryrd}
\usepackage{fontawesome}

@@

\usetikzlibrary{
  graphs,       % Graph *notation*
  graphdrawing, % Graph *layout*
  quotes,       % Quote syntax (e.g., "foo")
}

@@

\tikzset{graphs/graph style/.style={
 tree layout,
 level sep=1cm,
 nodes={draw, rounded corners=1mm},
 edges={draw, font={\small\ttfamily}}
}}
\tikzset{graphs/root style/.style={
 %draw=none,
 as={\textbullet$_{0}$}
}}
%\tikzset{graphs/every edge/.style={}

@@

This section introduces collaborative structure editing in Grove.
We begin in \autoref{sub:Program Representation} by covering how we represent expressions with holes as graphs,
\autoref{sub:Single User Edits} explores edits by a single user, Alice,
and \autoref{sub:Collaborative Structure Editing} explores edits by multiple users, Alice and Bob,
each of whom is editing their own instance of a program.

@@

@@

\subsubsection{Additions}
\label{sub:Additions}

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
TODO
\caption{User interface}
\label{fig:Addition-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{*}_{2}$" [> "root", > "1"'] -> {
   "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
   "$\texttt{var(y)}_{6}$" [> "right", > "5"']
  }
 }
};
\end{tikzpicture}
\caption{\autoref{fig:A simple tree with a hole-b} after adding \Vertex{6} and \Edge{5}}
\label{fig:Addition-b}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Addition}
\label{fig:Addition}
\end{figure}

TODO: Actions act at a cursor.  Each instance also has a cursor that it controls, but this is not part of the graph representation.

Alice starts with the code in \autoref{fig:A simple tree with a hole},
She moves her cursor to the hole in~\texttt{x * \hole} and enters the variable~\texttt{y}.
We model edits as sequences of additions and deletions of edges to the graph.
Alice's edit corresponds to creating a vertex for the variable reference~\texttt{y} and adding an edge
to that vertex from \Vertex{2}.
\autoref{fig:Addition-b} shows the result, where the new vertex has identifier~6.
This graph represents the expression \texttt{x * y}.

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{*}_{2}$" [> "root", > "1"'] -> {
   "$\texttt{x}_{4}$" [> "left", > "3"'],
   {}
  }
 }
};
\path (2cm,-2cm) graph[graph style] {
 y [as={$\texttt{y}_{6}$}]
};
\end{tikzpicture}
\caption{\autoref{fig:Addition-b} after deleting \Edge{5}}
\label{fig:Change-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{*}_{2}$" [> "RT$_1$"] -> {
   "$\texttt{x}_{4}$" [> "L$_3$"'],
   "$\texttt{z}_{8}$" [> "R$_7$"]
  }
 }
};
\path (2cm,-2cm) graph[graph style] {
 "$\texttt{y}_{6}$"
};
\end{tikzpicture}
\caption{\autoref{fig:Change-a} after adding \Vertex{8} and \Edge{7}}
\label{fig:Change-b}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Change}
\label{fig:Change}
\end{figure}

@@

\subsection{Collaborative Structure Editing}
\label{sub:Collaborative Structure Editing}

>>>>

%% Packages
\usepackage{booktabs}
\usepackage[rule=false]{subcaption}
\usepackage{graphicx}
\usepackage{semantic}
\let\colonapprox\undefined % Avoid redefinition error in `colonequals`
\let\colonsim\undefined % Avoid redefinition error in `colonequals`
\usepackage{colonequals}
\usepackage{mathpartir}
\usepackage{stmaryrd}
\usepackage{fontawesome}

@@

\usetikzlibrary{
  backgrounds,  % Provides "framed" and "gridded"
  graphs,       % Graph *notation*
  graphdrawing, % Graph *layout*
  quotes,       % Quote syntax (e.g., "foo")
}

@@

\tikzset{
  every picture/.style={framed, background rectangle/.style={draw=gray!50}},
}
\tikzset{graphs/graph style/.style={
  tree layout,
  level sep=1cm,
  nodes={draw, rounded corners=1mm},
  edges={draw, font={\small\ttfamily}},
  /tikz/every edge quotes/.style={
    %draw=gray!20,
    anchor=west,
    swap/.append code={
      \ifpgfarrowswap
        \pgfkeysalso{anchor=west}
      \else
        \pgfkeysalso{anchor=east}
      \fi}},
}}
\tikzset{graphs/root style/.style={
 %draw=none,
 as={\textbullet$_{0}$}
}}

@@

This section introduces collaborative structure editing in Grove.
We begin in \autoref{sub:Program Representation} by covering how we represent expressions with holes as graphs,
\autoref{sub:Single User Edits} explores edits by a single user, Alice,
and \autoref{sub:Multi-user Edits} explores edits by multiple users, Alice and Bob,
each of whom is editing their own instance of a program.

@@

In the remainder of this section,
we will consider user edits, starting
in this subsection with single-user
edits and continuing in \autoref{sub:Multi-user Edits}
with multi-user edits.

Each user edit that we will consider corresponds to one or more graph edits.
Graph edits can only add fresh edges and vertices or remove existing edges.

User edits act relative to a cursor. 
We do not model cursor sharing and the cursor does not appear in the underlying graph.
We discuss cursor representations and cursor sharing in more detail in \autoref{sub:Cursors}.

\subsubsection{Construction}
\label{sub:Construction}

\begin{figure}
\begin{tabular}{|p{0.49\linewidth}|p{0.49\linewidth}|}
%\centering
TODO
&
%\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{*}_{2}$" [> "Rt$_1$"] -> {
   "$\texttt{x}_{4}$" [> "L$_3$"'],
   "$\texttt{y}_{6}$" [> "R$_5$"]
  }
 }
};
\end{tikzpicture}
\\
\subcaption{User interface}
\label{fig:Construction-a}
 &
\subcaption{\autoref{fig:A simple tree with a hole-b} after adding \Vertex{6} and \Edge{5}}
\label{fig:Construction-b}
\\ \hline
\end{tabular}
\caption{Construction}
\label{fig:Construction}
\end{figure}

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}{0.49\columnwidth}
\centering
TODO
\caption{User interface}
\label{fig:Construction-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.49\columnwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{*}_{2}$" [> "Rt$_1$"] -> {
   "$\texttt{x}_{4}$" [> "L$_3$"'],
   "$\texttt{y}_{6}$" [> "R$_5$"]
  }
 }
};
\end{tikzpicture}
\caption{\autoref{fig:A simple tree with a hole-b} after adding \Vertex{6} and \Edge{5}}
\label{fig:Construction-b}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Construction}
\label{fig:Construction}
\end{figure}

TODO: "user edits" and "graph edits"

To start our examples, Alice starts with the code in \autoref{fig:A simple tree with a hole}.
She moves her cursor to the hole in~\texttt{x * \hole} and enters the variable~\texttt{y}.
We model edits as sequences of additions and deletions of edges to the graph.
Alice's edit corresponds to creating a vertex for the variable reference~\texttt{y} and adding an edge
to that vertex from \Vertex{2}.
\autoref{fig:Construction-b} shows the result, where the new vertex has identifier~6.
This graph represents the expression \texttt{x * y}.

@@

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}{0.49\linewidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{*}_{2}$" [> "root", > "1"'] -> {
   "$\texttt{x}_{4}$" [> "left", > "3"'],
   {}
  }
 }
};
\path (1cm,-2cm) graph[graph style] {
 y [as={$\texttt{y}_{6}$}]
};
\end{tikzpicture}
\caption{\autoref{fig:Construction-b} after deleting \Edge{5}}
\label{fig:Change-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.49\linewidth}
\centering
\begin{tikzpicture}[framed]
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{*}_{2}$" [> "RT$_1$"] -> {
   "$\texttt{x}_{4}$" [> "L$_3$"'],
   "$\texttt{z}_{8}$" [> "R$_7$"]
  }
 }
};
\path (1cm,-2cm) graph[graph style] {
 "$\texttt{y}_{6}$"
};
\end{tikzpicture}
\caption{\autoref{fig:Change-a} after adding \Vertex{8} and \Edge{7}}
\label{fig:Change-b}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Change}
\label{fig:Change}
\end{figure}

@@

\subsection{Multi-user Edits}
\label{sub:Multi-user Edits}
